---
title: "Diel Lake Carbon — V9 Model Analysis"
author: "John Ayers"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    embed-resources: true
    theme: flatly
execute:
  warning: false
  message: false
---

## Setup

```{r}
# Add user library to path if packages are installed there
user_lib <- file.path(Sys.getenv("R_LIBS_USER"))
if (dir.exists(user_lib)) .libPaths(c(user_lib, .libPaths()))

library(tidyverse)
library(readxl)
library(scales)
library(purrr)

df <- read_xlsx("Lakes\\Diel_v9_results_allseasons.xlsx") |>
  mutate(
    hour   = lubridate::hour(DateTime),
    Season = factor(Season, levels = c("Spring", "Summer", "Fall", "Winter"))
  )

season_colors <- c(
  Spring = "#4DAF4A",
  Summer = "#E41A1C",
  Fall   = "#FF7F00",
  Winter = "#377EB8"
)

theme_set(theme_bw(base_size = 12))
```

---

## Plot 1: δ¹³C Residuals by Hour of Day — Spring & Summer

Each point is one hourly observation. The smooth curve is a LOESS fit.

```{r}
#| fig-width: 9
#| fig-height: 4.5
#| fig-cap: "δ¹³C residuals (measured − modeled) vs. hour of day for Spring and Summer."

p1 <- df |>
  filter(Season %in% c("Spring", "Summer")) |>
  ggplot(aes(x = hour, y = d13C_residual, color = Season)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(size = 2, alpha = 0.75) +
  geom_smooth(method = "loess", span = 0.6, se = TRUE, linewidth = 1) +
  facet_wrap(~Season) +
  scale_x_continuous(breaks = seq(0, 23, 3),
                     labels = sprintf("%02d:00", seq(0, 23, 3))) +
  scale_color_manual(values = season_colors, guide = "none") +
  labs(
    title = expression(delta^{13}*"C Residuals by Hour of Day"),
    x     = "Hour of Day",
    y     = expression(delta^{13}*"C residual (‰)")
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

show(p1)
```

---

## Plot 2: δ¹³C-DIC Measured vs. Modeled — All Seasons

The dashed line is the 1:1 reference.

```{r}
#| fig-width: 6
#| fig-height: 5.5
#| fig-cap: "Measured vs. modeled δ¹³C-DIC for all four seasons combined."

range_vals <- range(c(df$d13C_DIC_measured, df$d13C_DIC_modeled), na.rm = TRUE)

p2 <- df |>
  ggplot(aes(x = d13C_DIC_modeled, y = d13C_DIC_measured, color = Season)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey40") +
  geom_point(size = 2.5, alpha = 0.8) +
  coord_fixed(xlim = range_vals, ylim = range_vals) +
  scale_color_manual(values = season_colors) +
  labs(
    title = expression(delta^{13}*"C-DIC: Measured vs. Modeled"),
    x     = expression(delta^{13}*"C-DIC modeled (‰)"),
    y     = expression(delta^{13}*"C-DIC measured (‰)"),
    color = "Season"
  )

p2
```

---

## Plot 3: Ca²⁺ vs. DIC by Season with SI~Calcite~ Color Gradient

Point color encodes the Calcite Saturation Index (SI > 0 = supersaturated; SI < 0 = undersaturated).

```{r}
#| fig-width: 10
#| fig-height: 8
#| fig-cap: "Ca²⁺ vs. DIC faceted by season. Color shows SI_Calcite; diverging scale centered at 0 (saturation boundary)."

p3 <- df |>
  ggplot(aes(x = DIC_mgL, y = Ca_measured_mgL, color = SI_Calcite)) +
  geom_point(size = 3, alpha = 0.85) +
  facet_wrap(~Season, scales = "free") +
  scale_color_gradient2(
    midpoint = 0,
    low      = "#2166AC",
    mid      = "#F7F7F7",
    high     = "#D6604D",
    name     = "SI Calcite",
    limits   = c(-1.1, 1.1)
  ) +
  labs(
    title = expression("Ca"^{"2+"}*" vs. DIC by Season with Calcite Saturation Index"),
    x     = "DIC (mg/L)",
    y     = expression("Ca"^{"2+"}*" (mg/L)")
  )

p3
```

---

## Plot 4: PAR vs. GPP~modeled~ by Season

```{r}
#| fig-width: 10
#| fig-height: 8
#| fig-cap: "PAR (W m⁻²) vs. modeled GPP (mg O₂ L⁻¹ h⁻¹) for each season."

p4 <- df |>
  ggplot(aes(x = PAR_Wm2, y = GPP_modeled_mgO2Lh, color = Season)) +
  geom_point(size = 2.5, alpha = 0.8) +
  geom_smooth(method = "loess", se = FALSE, linewidth = 1) +
  facet_wrap(~Season, scales = "free") +
  scale_color_manual(values = season_colors, guide = "none") +
  labs(
    title = "PAR vs. Modeled GPP by Season",
    x     = expression("PAR (W m"^{-2}*")"),
    y     = expression("GPP"[modeled]*" (mg O"[2]*" L"^{-1}*" h"^{-1}*")")
  )

p4
```

---

## Plot 5: Arrhenius Plot of ER~measured~

The Arrhenius equation: $\ln(ER) = \ln(A) - E_a / (R \cdot T)$, where $T$ is temperature in Kelvin, $R = 8.314$ J mol⁻¹ K⁻¹, and $E_a$ is the activation energy. ER values (O₂ units, negative by convention) are converted to absolute values before taking the natural log.

```{r}
#| fig-width: 8
#| fig-height: 6
#| fig-cap: "Arrhenius plot: 1/T (K⁻¹) vs. ln(|ER_measured|) with per-season linear regressions."

df_arr <- df |>
  filter(ER_observed_mgO2Lh != 0) |>
  mutate(
    T_inv = 1 / (Temperature_C + 273.15),
    ln_ER = log(abs(ER_observed_mgO2Lh))
  ) |>
  filter(is.finite(ln_ER), is.finite(T_inv))

# Fit lm per season and extract stats
arr_fits <- df_arr |>
  group_by(Season) |>
  nest() |>
  mutate(
    fit       = map(data, \(d) lm(ln_ER ~ T_inv, data = d)),
    slope     = map_dbl(fit, \(m) coef(m)[[2]]),
    intercept = map_dbl(fit, \(m) coef(m)[[1]]),
    r2        = map_dbl(fit, \(m) summary(m)$r.squared)
  ) |>
  select(-data, -fit)

# Build annotation labels, positioned at 80th/90th percentile of each season
arr_labels <- arr_fits |>
  mutate(
    label = sprintf("y = %.0fx %+.1f\nR² = %.3f", slope, intercept, r2),
    T_inv = map_dbl(Season, \(s) quantile(df_arr$T_inv[df_arr$Season == s], 0.80)),
    ln_ER = map_dbl(Season, \(s) quantile(df_arr$ln_ER[df_arr$Season == s], 0.90))
  )

p5 <- ggplot(df_arr, aes(x = T_inv, y = ln_ER, color = Season)) +
  geom_point(size = 2.5, alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  geom_label(
    data        = arr_labels,
    aes(label   = label),
    size        = 3,
    fill        = alpha("white", 0.8),
    label.size  = 0.3,
    show.legend = FALSE
  ) +
  scale_color_manual(values = season_colors) +
  scale_x_continuous(labels = label_scientific(digits = 3)) +
  labs(
    title = expression("Arrhenius Plot of Measured Ecosystem Respiration"),
    x     = expression("1/T (K"^{-1}*")"),
    y     = expression("ln(|ER"[measured]*"|)  [ln(mg O"[2]*" L"^{-1}*" h"^{-1}*")]"),
    color = "Season"
  )

p5
```

### Regression summary table

```{r}
arr_fits |>
  mutate(
    `Ea (kJ/mol)` = round(-slope * 8.314 / 1000, 1),
    slope         = round(slope, 0),
    intercept     = round(intercept, 2),
    r2            = round(r2, 4)
  ) |>
  rename(Slope = slope, Intercept = intercept, `R²` = r2) |>
  knitr::kable(caption = "Arrhenius regression results per season. Ea estimated from slope × (−R).")
```

---

## Save Plots

```{r}
if (!requireNamespace("svglite", quietly = TRUE))
  install.packages("svglite", repos = "https://cloud.r-project.org",
                   lib = Sys.getenv("R_LIBS_USER"))
library(svglite)

dir.create("plots", showWarnings = FALSE)

plot_specs <- list(
  list(plot = p1, name = "plot1_d13C_residuals_hourly",          w = 9,  h = 4.5),
  list(plot = p2, name = "plot2_d13C_DIC_measured_vs_modeled",   w = 6,  h = 5.5),
  list(plot = p3, name = "plot3_Ca_vs_DIC_SI_calcite",           w = 10, h = 8),
  list(plot = p4, name = "plot4_PAR_vs_GPP",                     w = 10, h = 8),
  list(plot = p5, name = "plot5_Arrhenius_ER_measured",          w = 8,  h = 6)
)

walk(plot_specs, \(p) {
  ggsave(file.path("plots", paste0(p$name, ".png")),
         plot = p$plot, width = p$w, height = p$h, dpi = 300, units = "in")
  ggsave(file.path("plots", paste0(p$name, ".svg")),
         plot = p$plot, width = p$w, height = p$h, units = "in",
         device = svglite)
})

message("Saved 5 plots × 2 formats (PNG + SVG) to plots/")
```
