---
title: "Lake Metabolism"
output: html_notebook
---

Adapting some of the metabolism models in the package "LakeMetabolizer":

Winslow L. A., Zwart J. A., Batt R. D., Dugan H. A., Woolway R. I., Corman J. R., Hanson P. C. and Read J. S. (2016) LakeMetabolizer: an R package for estimating lake metabolism from free-water oxygen using diverse statistical models. Inl. Waters 6, 622–636. Want instantaneous values of GPP and ER rather than just daily average values. Use sample data in LakeMetabolizer demo.

Some of the code is from the demo files and some from the package.

### Import data and define input values

```{r}
library(LakeMetabolizer)
data.path = system.file('extdata/', package="LakeMetabolizer")
sp.data = load.all.data('sparkling', data.path)
ts.data = sp.data$data #pull out just the timeseries data


#calculate U10 and add it back onto the original
u10 = wind.scale(ts.data)
ts.data = rmv.vars(ts.data, 'wnd', ignore.offset=TRUE) #drop old wind speed column
ts.data = merge(ts.data, u10) #merge new u10 into big dataset


#Now calculate k600 using the Cole method
k600.cole = k.cole(ts.data)

ts.data = merge(ts.data, k600.cole)

kgas = k600.2.kGAS(ts.data)
ts.data = rmv.vars(merge(kgas, ts.data), 'k600')

o2.sat = o2.at.sat(ts.data[,c('datetime','wtr_0')])

ts.data = merge(o2.sat, ts.data)
z.mix = ts.meta.depths(get.vars(ts.data, 'wtr'), seasonal=TRUE)
names(z.mix) = c('datetime','z.mix', 'bottom')

#set z.mix to bottom of lake when undefined
z.mix[z.mix$z.mix <=0 | is.na(z.mix$z.mix), 'z.mix'] = sp.data$metadata$maxdepth
ts.data = merge(ts.data, z.mix[,c('datetime','z.mix')])

```

### Try reporting instantaneous values. First define functions.

```{r}
calc.freq <- function (datetime) 
{
  freq <- round(Mode(1/diff(date2doy(datetime))))
}

Mode <- function (x) 
{
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

date2doy <- function (x) 
{
  stopifnot(any(grepl("^POSIX", class(x[1]))))
  day <- as.numeric(format(x, "%j"))
  pat <- quote("([0-9]{2}:){2}[0-9]{2}")
  midnight <- as.POSIXct(gsub(pat, "00:00:00", x), tz = "GMT")
  frac <- as.numeric(difftime(x, midnight, units = "days"))
  day + frac
}

OLS_inst <- function (datetime, do.obs, do.sat, k.gas, z.mix, irr, wtr, ...) 
{
  # complete.inputs(do.obs = do.obs, do.sat = do.sat, k.gas = k.gas, 
  #                 z.mix = z.mix, irr = irr, wtr = wtr, error = TRUE)
  nobs <- length(do.obs)
  mo.args <- list(...)
  if (any(z.mix <= 0)) {
    stop("z.mix must be greater than zero.")
  }
  if (any(wtr <= 0)) {
    stop("all wtr must be positive.")
  }
  # if ("datetime" %in% names(mo.args)) {
  #   datetime <- mo.args$datetime
  #   freq <- calc.freq(datetime)
  #   if (nobs != freq) {
  #     bad.date <- format.Date(datetime[1], format = "%Y-%m-%d")
  #     warning("number of observations on ", bad.date, " (", 
  #             nobs, ") ", "does not equal estimated sampling frequency", 
  #             " (", freq, ")", sep = "")
  #   }
  # }
  # else {
  #   warning("datetime not found, inferring sampling frequency from # of observations")
  #   freq <- nobs
  # }
  freq <- calc.freq(datetime)
  do.diff <- diff(do.obs)
  inst_flux <- (k.gas/freq) * (do.sat - do.obs)
  flux <- inst_flux[-nobs]
  noflux.do.diff <- do.diff - flux/z.mix[-nobs]
  lntemp <- log(wtr)
  mod <- lm(noflux.do.diff ~ irr[-nobs] + lntemp[-nobs] - 1)
  rho <- mod[[1]][2]
  iota <- mod[[1]][1]
  mod.matrix <- model.matrix(mod)
  gpp <- iota * mod.matrix[, 1]*freq
  resp <- rho * mod.matrix[, 2]*freq
  nep <- gpp + resp
  datetime <- datetime[-1]
  inst.results <- list(mod = mod, metab = data.frame(datetime = datetime, GPP = gpp, 
                                                R = resp, NEP = nep))
  return(inst.results)
}
```

### Now run ordinary least squares model

```{r}
ols.inst.res = OLS_inst(datetime = ts.data$datetime, 
                        do.obs = ts.data$doobs_0.5, do.sat = ts.data$do.sat,
                        k.gas = ts.data$k.gas, z.mix = ts.data$z.mix,
                        irr = ts.data$par, wtr = ts.data$wtr_0.5)
# Get error message when I run the following line
#write.csv(ols.inst.res$metab, 'LakeAnalyzer/metab.ols.inst.csv', row.names=FALSE)
ols.inst.res.df <- data.frame(ols.inst.res$metab)
plot(ols.inst.res.df$datetime, ols.inst.res.df$GPP)

```

### Bayesian metabolism model. First define functions

```{r}
bayes.makeModel <- function (k.gas) 
{
  if (!requireNamespace("R2jags")) {
    stop("metab.bayesian requires R2jags.\ninstall.packages('R2jags')\n #Also install JAGS (http://mcmc-jags.sourceforge.net/)")
  }
  finite.1oK <- is.finite(1/k.gas)
  choose.allK <- all(finite.1oK)
  choose.noK <- all(!finite.1oK)
  choose.bothK <- any(finite.1oK) & any(!finite.1oK)
  choice.mod <- c("allK", "noK", "bothK")[c(choose.allK, choose.noK, 
                                            choose.bothK)]
  jags.dir = system.file("jags", package = "LakeMetabolizer")
  switch(choice.mod, allK = {
    modfile = file.path(jags.dir, "bayes.mod.allk.JAGS")
    print("using allK model")
  }, noK = {
    modfile = file.path(jags.dir, "bayes.mod.nok.JAGS")
    print("using noK model")
  }, bothK = {
    modfile = file.path(jags.dir, "bayes.mod.bothk.JAGS")
    print("using bothK model")
  })
  return(modfile)
} 

bayesFit <- function (data, params, mf, tend = "median", ...) 
{
  bf.args <- list(...)
  jags.m <- R2jags::jags(data, NULL, parameters.to.save = params, 
                         mf)
  tF <- function(x, tend) {
    switch(tend, median = median(x), mean = mean(x), mode1 = unique(x)[which.max(tabulate(match(x, 
                                                                                                unique(x))))], mode2 = {
                                                                                                  xd <- density(x)
                                                                                                  xd$x[which.max(xd$y)]
                                                                                                })
  }
  simOut <- jags.m$BUGSoutput$sims.matrix
  ctSim <- apply(simOut, 2, tF, tend)
  sdSim <- apply(simOut, 2, sd)
  n.obs <- length(data$U[, 1])
  GPP <- mean(ctSim[1] * data$U[, 1], na.rm = TRUE) * n.obs
  R <- mean(ctSim[2] * data$U[, 2], na.rm = TRUE) * n.obs
  GPPsd <- sqrt(sum(sdSim[1]^2 * data$U[, 1]^2))
  Rsd <- sqrt(sum(sdSim[2]^2 * data$U[, 2]^2))
  NEPsd <- sqrt(GPPsd^2 + Rsd^2)
  return(list(model = jags.m, params = ctSim[1:2], metab.sd = matrix(c(GPPsd, 
                                                                       Rsd, NEPsd), nrow = 1, dimnames = list(NULL, c("GPPsd", 
                                                                                                                      "Rsd", "NEPsd"))), metab = matrix(c(GPP, R, GPP + R), 
                                                                                                                                                        nrow = 1, dimnames = list(NULL, c("GPP", "R", "NEP")))))
}

BAYES_inst <- function (datetime, do.obs, do.sat, k.gas, z.mix, irr, wtr, priors, ...) 
{
  # complete.inputs(do.obs = do.obs, do.sat = do.sat, k.gas = k.gas, 
  #                 z.mix = z.mix, irr = irr, wtr = wtr, error = TRUE)
  # if (any(z.mix <= 0)) {
  #   stop("z.mix must be greater than zero.")
  # }
  # if (any(wtr <= 0)) {
  #   stop("all wtr must be positive.")
  # }
  mb.args <- list(...)
  nobs <- length(do.obs)
  # if ("datetime" %in% names(mb.args)) {
  #   datetime <- mb.args$datetime
  #   freq <- calc.freq(datetime)
  #   if (nobs != freq) {
  #     bad.date <- format.Date(datetime[1], format = "%Y-%m-%d")
  #     warning("number of observations on ", bad.date, " (", 
  #             nobs, ") ", "does not equal estimated sampling frequency", 
  #             " (", freq, ")", sep = "")
  #   }
  # }
  # else {
  #   warning("datetime not found, inferring sampling frequency from # of observations")
  #   freq <- nobs
  # }
  freq <- calc.freq(datetime)
  if (!missing(priors)) {
    t.priors <- priors
    p.name.logic <- all(c(c("gppMu", "gppSig2", "rMu", "rSig2", 
                            "kSig2")) %in% names(t.priors))
    p.class.logic <- is.integer(t.priors) | is.numeric(t.priors)
    if (p.name.logic & p.class.logic) {
      priors <- t.priors
    }
    else {
      stop("supplied priors was not a (properly) named numeric/integer vector")
    }
  }
  else {
    priors <- c(gppMu = 0, gppSig2 = 1e+05, rMu = 0, rSig2 = 1e+05, 
                kSig2 = NA)
  }
  if (!all(c(is.numeric(do.obs), is.numeric(do.sat), is.numeric(k.gas), 
             is.numeric(z.mix), is.numeric(irr), is.numeric(wtr)))) {
    stop("All inputs to metab.bayes must be numeric vectors")
  }
  requireNamespace("R2jags")
  modfile <- bayes.makeModel(k.gas = (k.gas/freq))
  U <- matrix(NA, nrow = length(irr), ncol = 2)
  U[, 1] <- irr
  U[, 2] <- log(wtr)
  kP <- matrix(NA, nrow = length(k.gas), ncol = 2)
  kP[, 1] <- k.gas/freq
  if (is.na(priors["kSig2"])) {
    k0.logic <- !is.finite(1/kP[, 1])
    kP[, 2] <- sum(kP[, 1])/sum(!k0.logic) * 0.1
    kP[k0.logic, 2] <- 1e-09
  }
  else {
    kP[, 2] <- priors["kSig2"]
  }
  cP <- matrix(NA, nrow = 2, ncol = 2)
  cP[1, 1] <- priors["gppMu"]
  cP[1, 2] <- priors["gppSig2"]
  cP[2, 1] <- priors["rMu"]
  cP[2, 2] <- priors["rSig2"]
  data <- list(Y = do.obs, N = length(do.obs), U = U, kP = kP, 
               cP = cP, satO = do.sat, a0 = do.obs[1], Zmix = z.mix)
  params <- c("C", "sigmaV", "sigmaW")
  output <- bayesFit(data, params, mf = modfile)
  return(output)
}
```

### Run Bayesian model

```{r}
bayes.inst.res = BAYES_inst(datetime = ts.data$datetime, do.obs = ts.data$doobs_0.5, do.sat = ts.data$do.sat, k.gas = ts.data$k.gas,
                        z.mix = ts.data$z.mix, irr = ts.data$par, wtr = ts.data$wtr_0.5)
write.csv(bayes.inst.res$metab, 'LakeAnalyzer/metab.bayes.inst.csv', row.names=FALSE)

```
