---
title: "Lake Metabolism"
output:
  html_document:
    df_print: paged
---

Adapting some of the metabolism models in the package "LakeMetabolizer":

Winslow L. A., Zwart J. A., Batt R. D., Dugan H. A., Woolway R. I., Corman J. R., Hanson P. C. and Read J. S. (2016) LakeMetabolizer: an R package for estimating lake metabolism from free-water oxygen using diverse statistical models. Inl. Waters 6, 622–636. Want instantaneous values of GPP and ER rather than just daily average values. Use sample data in LakeMetabolizer demo.

Some of the code is from the demo files and some from the package.

### Import data and define input values

```{r}
library(LakeMetabolizer)
data.path = system.file('extdata/', package="LakeMetabolizer")
sp.data = load.all.data('sparkling', data.path)
ts.data = sp.data$data #pull out just the timeseries data


#calculate U10 and add it back onto the original
u10 = wind.scale(ts.data)
ts.data = rmv.vars(ts.data, 'wnd', ignore.offset=TRUE) #drop old wind speed column
ts.data = merge(ts.data, u10) #merge new u10 into big dataset


#Now calculate k600 using the Cole method
k600.cole = k.cole(ts.data)

ts.data = merge(ts.data, k600.cole)

kgas = k600.2.kGAS(ts.data)
ts.data = rmv.vars(merge(kgas, ts.data), 'k600')

o2.sat = o2.at.sat(ts.data[,c('datetime','wtr_0')])

ts.data = merge(o2.sat, ts.data)
z.mix = ts.meta.depths(get.vars(ts.data, 'wtr'), seasonal=TRUE)
names(z.mix) = c('datetime','z.mix', 'bottom')

#set z.mix to bottom of lake when undefined
z.mix[z.mix$z.mix <=0 | is.na(z.mix$z.mix), 'z.mix'] = sp.data$metadata$maxdepth
ts.data = merge(ts.data, z.mix[,c('datetime','z.mix')])

```

### Report instantaneous values. First define functions.
estimate GPP and R at each measurement frequency by multiplying the daily fitted iota and rho parameters by light and water temperature, respectively.  iota * light would give you an estimate of GPP for each time step in units of mg L-1 ∆time-1. If you multiply this by the number of observations that you have in a day (e.g., 10min sampling frequency would be 144 ∆time day-1) then you'd get units in mg L-1 day-1 for each time step.
iota is the primary production per quantum of light. rho is the average rate of respiration per natural log of water temperature. 

```{r}
calc.freq <- function (datetime) 
{
  freq <- round(Mode(1/diff(date2doy(datetime))))
}

Mode <- function (x) 
{
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

date2doy <- function (x) 
{
  stopifnot(any(grepl("^POSIX", class(x[1]))))
  day <- as.numeric(format(x, "%j"))
  pat <- quote("([0-9]{2}:){2}[0-9]{2}")
  midnight <- as.POSIXct(gsub(pat, "00:00:00", x), tz = "GMT")
  frac <- as.numeric(difftime(x, midnight, units = "days"))
  day + frac
}

OLS_inst <- function (datetime, do.obs, do.sat, k.gas, z.mix, irr, wtr, ...) 
{
  # complete.inputs(do.obs = do.obs, do.sat = do.sat, k.gas = k.gas, 
  #                 z.mix = z.mix, irr = irr, wtr = wtr, error = TRUE)
  nobs <- length(do.obs)
  mo.args <- list(...)
  if (any(z.mix <= 0)) {
    stop("z.mix must be greater than zero.")
  }
  if (any(wtr <= 0)) {
    stop("all wtr must be positive.")
  }
  # if ("datetime" %in% names(mo.args)) {
  #   datetime <- mo.args$datetime
  #   freq <- calc.freq(datetime)
  #   if (nobs != freq) {
  #     bad.date <- format.Date(datetime[1], format = "%Y-%m-%d")
  #     warning("number of observations on ", bad.date, " (", 
  #             nobs, ") ", "does not equal estimated sampling frequency", 
  #             " (", freq, ")", sep = "")
  #   }
  # }
  # else {
  #   warning("datetime not found, inferring sampling frequency from # of observations")
  #   freq <- nobs
  # }
  freq <- calc.freq(datetime)
  do.diff <- diff(do.obs)
  inst_flux <- (k.gas/freq) * (do.sat - do.obs)
  flux <- inst_flux[-nobs]
  noflux.do.diff <- do.diff - flux/z.mix[-nobs]
  lntemp <- log(wtr)
  mod <- lm(noflux.do.diff ~ irr[-nobs] + lntemp[-nobs] - 1)
  rho <- mod[[1]][2]
  iota <- mod[[1]][1]
  mod.matrix <- model.matrix(mod)
  gpp <- iota * mod.matrix[, 1]*freq
  resp <- rho * mod.matrix[, 2]*freq
  nep <- gpp + resp
  datetime <- datetime[-1]
  inst.results <- list(mod = mod, metab = data.frame(datetime = datetime, GPP = gpp, R = resp, NEP = nep))
  return(inst.results)
}
```

### Now run ordinary least squares model
GPP, ER and NEP reported in mg L-1 day-1

```{r}
ols.inst.res = OLS_inst(datetime = ts.data$datetime,
                        do.obs = ts.data$doobs_0.5, do.sat = ts.data$do.sat,
                        k.gas = ts.data$k.gas, z.mix = ts.data$z.mix,
                        irr = ts.data$par, wtr = ts.data$wtr_0.5)
ols.inst.res.df <- data.frame(ols.inst.res$metab)
plot(ols.inst.res.df$datetime, ols.inst.res.df$GPP)
plot(ols.inst.res.df$datetime, ols.inst.res.df$R)
plot(ols.inst.res.df$datetime, ols.inst.res.df$NEP)
# Get error message when I run the following line
write.csv(ols.inst.res.df, 'metab.ols.inst.csv', row.names=FALSE)

```

```{r}
#Bookkeep
# add column for day (1) and night (0)
ts.data$day_night = as.integer(is.day(datetimes=ts.data$datetime, lat=sp.data$metadata$latitude))

book.res = metab(ts.data, method='bookkeep',
                 wtr.name='wtr_0.5', do.obs.name='doobs_0.5', irr.name='day_night')

#Bring year and DOY together to get R datetime again
ols.inst.res$datetime = ISOdate(ols.inst.res$year, 1, 1) + book.res$doy*3600*24 - 3600*24

#summary stats
res = ols.inst.res.df
cat('metab.ols GPP:', mean(res[,2]), ' R:', mean(res[,3]), 'NEP:', mean(res[,4]), '\n')

#impossible values removed
cat('Averages after removal of impossible GPP and R values\n')
cat('metab.ols GPP:', mean(res[res[,2]>0,2]), ' R:', mean(res[res[,3]<0,3]), 'NEP:', mean(res[res[,2]>0,2]) + mean(res[res[,3]<0,3]), '\n')
```

